### 2018-05-07 doc/paper.tex
Large-scale changes.
Added a few lemmas with actual bounds and re-organized so
that the proof is mostly done, just needs a few blanks to
be filled in.

Throughout this paper we assume that the computation in floating point
arithmetic obeys the model
\begin{equation}
  a \star b = \fl{a \circ b} = (a \circ b)(1 + \eps_1) =
  (a \circ b) / (1 + \eps_2)
\end{equation}
where \(\star \in \left\{\oplus, \ominus, \otimes, \oslash\right\}\), \(\circ
\in \left\{+, -, \times, \div\right\}\) and \(\left|\eps_1\right|,
\left|\eps_2\right| \leq \mach\). The symbol \(\mach\) is the unit round-off
and \(\star\) represents the floating point computation, e.g.
\(a \oplus b = \fl{a + b}\). (For IEEE-754 floating point double precision,
\(\mach = 2^{-53}\).) We denote the computed result of
\(\alpha \in \mathbf{R}\) in floating point arithmetic by
\(\widehat{\alpha}\) or \(\fl{\alpha}\) and \(\mathbf{F}\) denotes the set of
all floating point numbers (see \cite{Higham2002} for more details).
Following \cite{Higham2002}, we will use the following classic properties in
error analysis.

\begin{enumerate}
  \item If \(\delta_i \leq \mach\), \(\rho_i = \pm 1\), then
      \(\prod_{i = 1}^n (1 + \delta_i)^{\rho_i} = 1 + \theta_n\),
  \item \(\left|\theta_n\right| \leq \gamma_n \coloneqq
      n \mach / (1 - n \mach)\),
  \item \((1 + \theta_k)(1 + \theta_j) \leq 1 + \theta_{k + j}\),
  \item \(\gamma_k + \gamma_j + \gamma_k \gamma_j \leq \gamma_{k + j}\),
  \item \((1 + \mach)^j \leq 1 / (1 - j \mach)\).
\end{enumerate}

@@

\begin{theorem}[\cite{Ogita2005}]\label{thm:eft}
For \(a, b \in \mathbf{F}\) and \(P, \pi, S, \sigma \in \mathbf{F}\),
\texttt{TwoSum} and \texttt{TwoProd} satisfy
\begin{itemize}
  \item \(\left[S, \sigma\right] = \mathtt{TwoSum}(a, b)\), \(x = \fl{a + b}\),
      \(S + \sigma = a + b\), \(\sigma \leq \mach \left|S\right|\),
      \(\sigma \leq \mach \left|a + b\right|\)
  \item \(\left[P, \pi\right] = \mathtt{TwoProd}(a, b)\),
      \(P = \fl{a \times b}\), \(P + \pi = a \times b\),
      \(\pi \leq \mach \left|P\right|\),
      \(\pi \leq \mach \left|a \times b\right|\).
\end{itemize}
The letters \(\sigma\) and \(\pi\) are used to indicate that the
errors came from sum and product, respectively. See the appendix
in Section~\ref{sec:appendix-eft} for implementation details.
\end{theorem}

@@

The relative condition number of the evaluation of \(p(s) = \sum_{j = 0}^n
b_j B_{j, n}(s)\) in Bernstein form used in this paper is (see
\cite{Mainar1999}, \cite{Farouki1987}):
\begin{equation}
\cond{p, s} = \frac{\widetilde{p}(s)}{
  \left|p(s)\right|} = \frac{\sum_j \left|b_j\right| B_{j, n}(s)}{
  \left|p(s)\right|},
\end{equation}
where \(B_{j, n} \geq 0\).

@@

\begin{table}
  \centering
  \begin{tabular}{>{$}c<{$} >{$}c<{$} >{$}c<{$} >{$}c<{$} >{$}c<{$} >{$}c<{$}}
    \toprule
    k & j & \widehat{b}_j^{(k)} & \cdb{1}_j^{(k)} & \db{1}_j^{(k)} - \cdb{1}_j^{(k)} \\
    \midrule
    3 & 0 & 0.125 - 1.75 (1001 \mach) - 0.25 \mach & 0.25\mach & 0 \\
    3 & 1 & -0.125 + 1.25(1001 \mach) + 0.25 \mach & -0.25\mach & 0 \\
    3 & 2 & 0.125 - 0.75 (1001 \mach) & 0 & 0 \\
    3 & 3 & -0.125 + 0.25 (1001 \mach) & 0 & 0 \\
    \midrule
    2 & 0 & -0.5 (1001 \mach) & 3 (1001 \mach)^2 & 0 \\
    2 & 1 & 0.5(1001 \mach) + 0.125 \mach & -0.125\mach - 2 (1001 \mach)^2 & 0 \\
    2 & 2 & -0.5 (1001 \mach) & (1001 \mach)^2 & 0 \\
    \midrule
    1 & 0 & 0.0625\mach + (1001 \mach)^2 + 239\mach^2 & -0.0625\mach + 0.5  (1001 \mach)^2 - 239 \mach^2 & -5 (1001\mach)^3 \\
    1 & 1 & 0.0625\mach - (1001 \mach)^2 - 239\mach^2 & -0.0625\mach - 0.5  (1001 \mach)^2 + 239 \mach^2 & 3 (1001\mach)^3 \\
    \midrule
    0 & 0 & 0.0625 \mach & -0.0625 \mach & -4 (1001 \mach)^3 + 8 (1001 \mach)^4 \\
    \bottomrule
  \end{tabular}
  \captionsetup{width=.75\linewidth}
  \caption{Terms computed by \texttt{CompDeCasteljau} when evaluating \\
    \(p(s) = (2s - 1)^3 (s - 1)\) at the point
    \(s = \frac{1}{2} + 1001 \mach\)}
  \label{tab:exact-computation}
\end{table}

@@

  \begin{algorithmic}
    \Function{\(\mathtt{res} = \mathtt{CompDeCasteljauK}\)}{$b, s, K$}
      \State \(n = \texttt{length}(b) - 1\)
      \State \(\left[\widehat{r}, \rho\right] = \mathtt{TwoSum}(1, -s)\)
      \\
      \For{\(j = 0, \ldots, n\)}
        \State \(\widehat{b}_j^{(n)} = b_j\)
        \For{\(F = 1, \ldots, K - 1\)}
          \State \(\cdb{F}_j^{(n)} = 0\)
        \EndFor
      \EndFor
      \\
      \For{\(k = n - 1, \ldots, 0\)}
        \For{\(j = 0, \ldots, k\)}
          \State \(\left[P_1, \pi_1\right] = \mathtt{TwoProd}\left(
              \widehat{r}, \widehat{b}_j^{(k + 1)}\right)\)
          \State \(\left[P_2, \pi_2\right] = \mathtt{TwoProd}\left(
              s, \widehat{b}_{j + 1}^{(k + 1)}\right)\)
          \State \(\left[\widehat{b}_j^{(k)}, \sigma_3\right] =
              \mathtt{TwoSum}(P_1, P_2)\)
          \\
          \State \(e = \left[\pi_1, \pi_2, \sigma_3\right]\)
          \State \(\delta b = \widehat{b}_j^{(k + 1)}\)
          \\
          \For{\(F = 1, \ldots, K - 2\)}
            \State \(\left[\eta, \widehat{\ell}\right] =
                \mathtt{LocalErrorEFT}(e, \rho, \delta b)\)
            \State \(M = \texttt{length}(\eta)\)
            \\
            \State \(\left[P_1, \eta_{M + 1}\right] = \mathtt{TwoProd}\left(
                s, \cdb{F}_{j + 1}^{(k + 1)}\right)\)
            \State \(\left[S_2, \eta_{M + 2}\right] =
                \mathtt{TwoSum}\left(\widehat{\ell}, P_1\right)\)
            \State \(\left[P_3, \eta_{M + 3}\right] = \mathtt{TwoProd}\left(
                \widehat{r}, \cdb{F}_j^{(k + 1)}\right)\)
            \State \(\left[\cdb{F}_j^{(k)}, \eta_{M + 4}\right]
                = \mathtt{TwoSum}\left(S_2, P_3\right)\)
            \\
            \State \(e = \eta\)
            \State \(\delta b = \cdb{F}_j^{(k + 1)}\)
          \EndFor
          \\
          \State \(\widehat{\ell} =
                \mathtt{LocalError}(e, \rho, \delta b)\)
          \State \(\cdb{K - 1}_j^{(k)} =
              \widehat{\ell} \oplus
              \left(s \otimes \cdb{K - 1}_{j + 1}^{(k + 1)}
              \right) \oplus
              \left(\widehat{r} \otimes
              \cdb{K - 1}_j^{(k + 1)}\right)\)
        \EndFor
      \EndFor
      \\
      \State \(\mathtt{res} = \widehat{b}_0^{(0)}\)
      \For{\(F = 1, \ldots, K - 1\)}
        \State \(\mathtt{res} = \mathtt{res} \oplus
            \cdb{F}_0^{(0)}\)
      \EndFor
    \EndFunction
  \end{algorithmic}
\end{breakablealgorithm}

\noindent Noting that \(\ell_{F, j}\) contains \(5F - 1\) terms, one can
show that \texttt{CompDeCasteljauK} (Algorithm~\ref{alg:k-comp-de-casteljau})
requires
\begin{equation}
(15K^2 + 11K - 34)T_n + K + 5 =
\mathcal{O}\left(n^2 K^2\right)
\end{equation}
flops to evaluate a degree \(n\) polynomial, where \(T_n\) is the
\(n\)th triangular number. As a comparison, the non-compensated form of
de Casteljau requires \(3 T_n + 1\) flops. In total this will require
\((3K - 4)T_n\) uses of \texttt{TwoProd}. On hardware that supports
FMA, \texttt{TwoProdFMA} (Algorithm~\ref{alg:two-prod-fma}) can be used
instead, lowering the flop count by \(15(3K - 4)T_n\).

@@

\begin{lemma}[\cite{Jiang2010}]
For \(p(s) = \sum_{j = 0}^n b_j B_{j, n}(s)\) a polynomial in Bernstein form
and \(s \in \left[0, 1\right]\), the terms produced by the de Casteljau
algorithm in exact arithmetic satisfy:
\begin{equation}
  \left|b_0^{(0)}\right| \leq \cdots \leq
  \sum_{j = 0}^k \left|b_j^{(k)}\right| B_{j, k}(s) \leq \cdots \leq
  \sum_{j = 0}^n \left|b_j\right| B_{j, n}(s) = \widetilde{p}(s).
\end{equation}
\end{lemma}

@@

\begin{theorem}
  If no underflow occurs, \(n \geq 2\) and \(s \in \left[0, 1\right]\)
  \begin{equation}
    \frac{\left|p(s) - \mathtt{CompDeCasteljau}(p, s, K)\right|}{
      \left|p(s)\right|} \leq \left[\mach + \mathcal{O}\left(\mach^2\right)
      \right] +
      \left[\left(\alpha(n) \mach\right)^K + \mathcal{O}\left(
      \mach^{K + 1}\right)\right] \cond{p, s}.
  \end{equation}
\end{theorem}

\begin{proof}
Note that
\begin{equation}
p(s) = b_0^{(0)} = \widehat{b}_0^{(0)} + \db{1}_0^{(0)}
= \cdots
= \widehat{b}_0^{(0)} + \cdb{1}_0^{(0)} + \cdots
+ \cdb{K - 1}_0^{(0)} + \db{K}_0^{(0)}
\end{equation}
and
\begin{equation}
\mathtt{CompDeCasteljau}(p, s, K) = \widehat{b}_0^{(0)} \oplus
\cdb{1}_0^{(0)} \oplus \cdots \oplus \cdb{K - 1}_0^{(0)}.
\end{equation}
It will be crucial for us to bound \(\db{K}_0^{(0)}\).
In addition, note that
\begin{align}
&\widehat{b}_0^{(0)} \oplus \cdb{1}_0^{(0)} \oplus \cdots \oplus
\cdb{K - 1}_0^{(0)} \\
&= \widehat{b}_0^{(0)} \left(1 + \theta_{K - 1}\right) +
\cdb{1}_0^{(0)} \left(1 + \theta_{K - 1}\right) + \cdots
+ \cdb{K - 1}_0^{(0)} \left(1 + \theta_1\right).
\end{align}
We'll use these two quantities to prove a bound.

First, note that \(\widehat{r} = (1 - s)(1 + \theta_1)\). Due to this,
we put the \(\widehat{r}\) term at the end of each update step to reduce
the amount of round-off:
\begin{align}
  \cdb{F}_j^{(k)} &=
  \widehat{\ell}_{F, j}^{(k)} \oplus
  \left(s \otimes \cdb{F}_{j + 1}^{(k + 1)}\right) \oplus
  \left(\widehat{r} \otimes \cdb{F}_j^{(k + 1)}\right) \\
&= (1 - s) \cdot \cdb{F}_j^{(k + 1)}(1 + \theta_3) +
  s \cdot \cdb{F}_{j + 1}^{(k + 1)}(1 + \theta_3) +
  \widehat{\ell}_{F, j}^{(k)} (1 + \theta_2) \\
\Longrightarrow \cdb{F}^{(k)} &=
  U_{k + 1}(s) \cdb{F}^{(k + 1)}(1 + \theta_3) +
  \widehat{\ell}_{F}^{(k)} (1 + \theta_2).
\end{align}
This is quite similar to the relationship satisfied in
exact arithmetic:
\begin{equation}
\db{F}^{(k)} = U_{k + 1}(s) \db{F}^{(k + 1)} + \ell_{F}^{(k)}.
\end{equation}

\end{proof}

>>>>

\subsection{Floating Point and Forward Error Analysis}

Throughout this paper we assume that the computation in floating point
arithmetic obeys the model
\begin{equation}
  a \star b = \fl{a \circ b} = (a \circ b)(1 + \delta_1) =
  (a \circ b) / (1 + \delta_2)
\end{equation}
where \(\star \in \left\{\oplus, \ominus, \otimes, \oslash\right\}\), \(\circ
\in \left\{+, -, \times, \div\right\}\) and \(\left|\delta_1\right|,
\left|\delta_2\right| \leq \mach\). The symbol \(\mach\) is the unit round-off
and \(\star\) represents the floating point computation, e.g.
\(a \oplus b = \fl{a + b}\). (For IEEE-754 floating point double precision,
\(\mach = 2^{-53}\).) We denote the computed result of
\(\alpha \in \mathbf{R}\) in floating point arithmetic by
\(\widehat{\alpha}\) or \(\fl{\alpha}\) and \(\mathbf{F}\) denotes the set of
all floating point numbers (see \cite{Higham2002} for more details).
Following \cite{Higham2002}, we will use the following classic properties in
error analysis.

\begin{enumerate}
  \item If \(\delta_i \leq \mach\), \(\rho_i = \pm 1\), then
      \(\prod_{i = 1}^n (1 + \delta_i)^{\rho_i} = 1 + \theta_n\),
  \item \(\left|\theta_n\right| \leq \gamma_n \coloneqq
      n \mach / (1 - n \mach)\),
  \item \((1 + \theta_k)(1 + \theta_j) \leq 1 + \theta_{k + j}\),
  \item \(\gamma_k + \gamma_j + \gamma_k \gamma_j \leq \gamma_{k + j}
    \Longleftrightarrow (1 + \gamma_k)(1 + \gamma_j) \leq 1 + \gamma_{k + j}\),
  \item \((1 + \mach)^j \leq 1 / (1 - j \mach) \Longleftrightarrow
  (1 + \mach)^j - 1 \leq \gamma_j\).
\end{enumerate}

\subsection{Error-Free Transformation}

@@

\begin{theorem}[\cite{Ogita2005}]\label{thm:eft}
For \(a, b \in \mathbf{F}\) and \(P, \pi, S, \sigma \in \mathbf{F}\),
\texttt{TwoSum} and \texttt{TwoProd} satisfy
\begin{itemize}
  \item \(\left[S, \sigma\right] = \mathtt{TwoSum}(a, b)\), \(x = \fl{a + b}\),
      \(S + \sigma = a + b\), \(\sigma \leq \mach \left|S\right|\),
      \(\sigma \leq \mach \left|a + b\right|\)
  \item \(\left[P, \pi\right] = \mathtt{TwoProd}(a, b)\),
      \(P = \fl{a \times b}\), \(P + \pi = a \times b\),
      \(\pi \leq \mach \left|P\right|\),
      \(\pi \leq \mach \left|a \times b\right|\).
\end{itemize}
The letters \(\sigma\) and \(\pi\) are used to indicate that the
errors came from sum and product, respectively. See the appendix
(Section~\ref{sec:appendix-algo}) for implementation details.
\end{theorem}

\subsection{de Casteljau Algorithm}

@@

The relative condition number of the evaluation of \(p(s) = \sum_{j = 0}^n
b_j B_{j, n}(s)\) in Bernstein form used in this paper is (see
\cite{Mainar1999}, \cite{Farouki1987}):
\begin{equation}
\cond{p, s} = \frac{\widetilde{p}(s)}{\left|p(s)\right|},
\end{equation}
where \(B_{j, n}(s) \geq 0\) and
\(\widetilde{p}(s) \coloneqq \sum_{j = 0}^n \left|b_j\right| B_{j, n}(s)\).

To be able to express the algorithm in matrix form, we define
the vectors
\begin{equation}
b^{(k)} = \left[\begin{array}{c c c} b_0^{(k)} & \cdots &
b_k^{(k)}\end{array}\right]^T, \quad
\widehat{b}^{(k)} = \left[\begin{array}{c c c} \widehat{b}_0^{(k)} & \cdots &
    \widehat{b}_k^{(k)}\end{array}\right]^T
\end{equation}
and the reduction matrices:
\begin{equation}
U_k = U_k(s) = \left[\begin{array}{c c c c c c}
    1 - s  & s      & 0      & \cdots & \cdots & 0      \\
    0      & 1 - s  & s      & \ddots &        & \vdots \\
    \vdots & \ddots & \ddots & \ddots & \ddots & \vdots \\
    \vdots &        & \ddots & \ddots & \ddots & 0 \\
    0      & \cdots & \cdots & 0      & 1 - s  & s
\end{array}\right] \in \mathbf{R}^{k \times (k + 1)}.
\end{equation}
With this, we can express (\cite{Mainar1999}) the de Casteljau algorithm as
\begin{equation}\label{matrix-de-casteljau}
b^{(k)} = U_{k + 1} b^{(k + 1)}
\Longrightarrow b^{(0)} = U_1 \cdots U_n b^{(n)}.
\end{equation}

In general, for a sequence \(v_0, \ldots, v_n\) we'll use the notation
\(v = \left[\begin{array}{c c c} v_0 & \cdots &
    v_n\end{array}\right]^T\) to indicate the vector containing all of
the values.

@@

\begin{table}
  %% See: ``scripts/verify_table.py``.
  \centering
  \begin{tabular}{>{$}c<{$} >{$}c<{$} >{$}c<{$} >{$}c<{$} >{$}c<{$} >{$}c<{$}}
    \toprule
    k & j & \widehat{b}_j^{(k)} & \cdb{1}_j^{(k)} & \db{1}_j^{(k)} - \cdb{1}_j^{(k)} \\
    \midrule
    3 & 0 & 0.125 - 1.75 (1001 \mach) - 0.25 \mach & 0.25\mach & 0 \\
    3 & 1 & -0.125 + 1.25(1001 \mach) + 0.25 \mach & -0.25\mach & 0 \\
    3 & 2 & 0.125 - 0.75 (1001 \mach) & 0 & 0 \\
    3 & 3 & -0.125 + 0.25 (1001 \mach) & 0 & 0 \\
    \midrule
    2 & 0 & -0.5 (1001 \mach) & 3 (1001 \mach)^2 & 0 \\
    2 & 1 & 0.5(1001 \mach) + 0.125 \mach & -0.125\mach - 2 (1001 \mach)^2 & 0 \\
    2 & 2 & -0.5 (1001 \mach) & (1001 \mach)^2 & 0 \\
    \midrule
    1 & 0 & 0.0625\mach + (1001 \mach)^2 + 239\mach^2 & -0.0625\mach + 0.5  (1001 \mach)^2 - 239 \mach^2 & -5 (1001\mach)^3 \\
    1 & 1 & 0.0625\mach - (1001 \mach)^2 - 239\mach^2 & -0.0625\mach - 0.5  (1001 \mach)^2 + 239 \mach^2 & 3 (1001\mach)^3 \\
    \midrule
    0 & 0 & 0.0625 \mach & -0.0625 \mach & -4 (1001 \mach)^3 + 8 (1001 \mach)^4 \\
    \bottomrule
  \end{tabular}
  \captionsetup{width=.75\linewidth}
  \caption{Terms computed by \texttt{CompDeCasteljau} when evaluating \\
    \(p(s) = (2s - 1)^3 (s - 1)\) at the point
    \(s = \frac{1}{2} + 1001 \mach\)}
  \label{tab:exact-computation}
\end{table}

@@

  \begin{algorithmic}
    \Function{\(\mathtt{res} = \mathtt{CompDeCasteljauK}\)}{$b, s, K$}
      \State \(n = \texttt{length}(b) - 1\)
      \State \(\left[\widehat{r}, \rho\right] = \mathtt{TwoSum}(1, -s)\)
      \\
      \For{\(j = 0, \ldots, n\)}
        \State \(\widehat{b}_j^{(n)} = b_j\)
        \For{\(F = 1, \ldots, K - 1\)}
          \State \(\cdb{F}_j^{(n)} = 0\)
        \EndFor
      \EndFor
      \\
      \For{\(k = n - 1, \ldots, 0\)}
        \For{\(j = 0, \ldots, k\)}
          \State \(\left[P_1, \pi_1\right] = \mathtt{TwoProd}\left(
              \widehat{r}, \widehat{b}_j^{(k + 1)}\right)\)
          \State \(\left[P_2, \pi_2\right] = \mathtt{TwoProd}\left(
              s, \widehat{b}_{j + 1}^{(k + 1)}\right)\)
          \State \(\left[\widehat{b}_j^{(k)}, \sigma_3\right] =
              \mathtt{TwoSum}(P_1, P_2)\)
          \\
          \State \(e = \left[\pi_1, \pi_2, \sigma_3\right]\)
          \State \(\delta b = \widehat{b}_j^{(k + 1)}\)
          \\
          \For{\(F = 1, \ldots, K - 2\)}
            \State \(\left[\eta, \widehat{\ell}\right] =
                \mathtt{LocalErrorEFT}(e, \rho, \delta b)\)
            \State \(L = \texttt{length}(\eta)\)
            \\
            \State \(\left[P_1, \eta_{L + 1}\right] = \mathtt{TwoProd}\left(
                s, \cdb{F}_{j + 1}^{(k + 1)}\right)\)
            \State \(\left[S_2, \eta_{L + 2}\right] =
                \mathtt{TwoSum}\left(\widehat{\ell}, P_1\right)\)
            \State \(\left[P_3, \eta_{L + 3}\right] = \mathtt{TwoProd}\left(
                \widehat{r}, \cdb{F}_j^{(k + 1)}\right)\)
            \State \(\left[\cdb{F}_j^{(k)}, \eta_{L + 4}\right]
                = \mathtt{TwoSum}\left(S_2, P_3\right)\)
            \\
            \State \(e = \eta\)
            \State \(\delta b = \cdb{F}_j^{(k + 1)}\)
          \EndFor
          \\
          \State \(\widehat{\ell} =
                \mathtt{LocalError}(e, \rho, \delta b)\)
          \State \(\cdb{K - 1}_j^{(k)} =
              \widehat{\ell} \oplus
              \left(s \otimes \cdb{K - 1}_{j + 1}^{(k + 1)}
              \right) \oplus
              \left(\widehat{r} \otimes
              \cdb{K - 1}_j^{(k + 1)}\right)\)
        \EndFor
      \EndFor
      \\
      \State \(\mathtt{res} = \mathtt{SumK}\left(\left[
        \widehat{b}_0^{(0)}, \ldots, \cdb{K - 1}_0^{(0)}\right], K\right)\)
    \EndFunction
  \end{algorithmic}
\end{breakablealgorithm}

\noindent Noting that \(\ell_{F, j}\) contains \(5F - 1\) terms, one can
show that \texttt{CompDeCasteljauK} (Algorithm~\ref{alg:k-comp-de-casteljau})
requires
\begin{equation}
(15K^2 + 11K - 34)T_n + 6K^2 - 11K + 11 =
\mathcal{O}\left(n^2 K^2\right)
\end{equation}
flops to evaluate a degree \(n\) polynomial, where \(T_n\) is the
\(n\)th triangular number. As a comparison, the non-compensated form of
de Casteljau requires \(3 T_n + 1\) flops. In total this will require
\((3K - 4)T_n\) uses of \texttt{TwoProd}. On hardware that supports
FMA, \texttt{TwoProdFMA} (Algorithm~\ref{alg:two-prod-fma}) can be used
instead, lowering the flop count by \(15(3K - 4)T_n\). Another way
to lower the total flop count is to just use
\(\widehat{b}_0^{(0)} \oplus \cdots \oplus \cdb{K - 1}_0^{(0)}\)
instead of \texttt{SumK}. This will reduce the total by
\(6(K - 1)^2\) flops, however the results produced are (empirically)
identical. This makes sense: the whole point of \texttt{SumK}
is to filter errors in a summation so that the final operation produces
a sum of the form \(v_1 \oplus \cdots \oplus v_K\) where each
term is smaller than the previous by a factor of \(\mach\). This
property is already satisfied for the \(\cdb{F}_0^{(0)}\) so in
practice the \(K\)-compensated summation is likely not needed.

@@

\begin{theorem}[\cite{Ogita2005}]\label{thm:sum-k}
A summation can be computed (\texttt{SumK}, Algorithm~\ref{alg:sum-k})
with results that are as accurate as if computed in \(K\) times the
working precision. When computed this way, the result satisfies:
\begin{equation}
\left|\mathtt{SumK}(v, K) - \sum_{j = 1}^n v_j\right| \leq
\left(\mach + 3 \gamma_{n - 1}^2\right) \left|\sum_{j = 1}^n v_j\right| +
\gamma_{2n - 2}^K \sum_{j = 1}^n \left|v_j\right|.
\end{equation}
\end{theorem}

@@

To enable a bound on the \(K\)-order error \(\db{K}^{(0)}_0\), it's necessary
to understand the difference between the exact local errors \(\ell_{F, j}\)
and the computed equivalents \(\widehat{\ell}_{F, j}\). To do this, we define
\begin{equation}
\widetilde{\ell}_{F, j} \coloneqq \left|e_1\right| +
 + \cdots + \left|e_{5F - 2}\right| + \left|\rho \cdot
\cdb{F - 1}_j^{(k + 1)}\right|.
\end{equation}

\begin{lemma}\label{lemma:ell-tilde}
The local errors \(\widetilde{\ell}_{F, j}\) satisfy:
\begin{align}
\widetilde{\ell}_{1, j}^{(k)} &\leq
  \gamma_3 \left(
  (1 - s) \left|\widehat{b}_j^{(k + 1)}\right| +
  s \left|\widehat{b}_{j + 1}^{(k + 1)}\right|\right) \\
\widetilde{\ell}_{F + 1, j}^{(k)} &\leq
  \gamma_3 \left(
  (1 - s) \left|\cdb{F}_j^{(k + 1)}\right| +
  s \left|\cdb{F}_{j + 1}^{(k + 1)}\right|\right) +
  \gamma_{5F} \cdot \widetilde{\ell}_{F, j}^{(k)}
  \text{ for } F \geq 1.
\end{align}
\end{lemma}

As we'll see soon (Lemma~\ref{lemma:k-order}), putting a bound on
sums of the form \(\sum_{j = 0}^k \ell_{F, j}^{(k)} B_{j, k}(s)\) will
be necessary to get an overall bound on the relative error for
\texttt{CompDeCasteljauK}. So we define
\(L_{F, k} \coloneqq \sum_{j = 0}^k \ell_{F, j}^{(k)} B_{j, k}(s)\) and the
related
\(D_{0, k} \coloneqq \sum_{j = 0}^k \left|\widehat{b}_j^{(k)}\right|
B_{j, k}(s)\) and
\(D_{F, k} \coloneqq \sum_{j = 0}^k \left|\cdb{F}_j^{(k)}\right| B_{j, k}(s)\).

\begin{lemma}\label{lemma:L-and-D-bounds}
For \(s \in \left[0, 1\right]\), the Bernstein-type sums defined above
satisfy the following bounds:
\begin{align}
D_{0, n - k} &\leq \left(1 + \gamma_{3k}\right) \cdot \widetilde{p}(s) \\
L_{1, n - k} &\leq \gamma_{3k} \cdot \widetilde{p}(s) \\
D_{F + 1, n - k} &\leq \binom{k + F}{F + 1} \gamma_3 \gamma_{5} \gamma_{10}
  \cdots \gamma_{5F} \cdot \left(1 + \gamma_{3k + 5F + 2}\right)
  \cdot \widetilde{p}(s) \\
L_{F + 2, n - k} &\leq \binom{k + F}{F + 1} \gamma_3 \gamma_{5} \gamma_{10}
  \cdots \gamma_{5F} \cdot \gamma_{3k + 5F + 2} \cdot \widetilde{p}(s).
\label{L-sum-bound}
\end{align}
In particular, this means that
\(D_{F, k} = \mathcal{O}\left(\mach^F\right) \cdot \widetilde{p}(s)\) and
\(L_{F, k} = \mathcal{O}\left(\mach^F\right) \cdot \widetilde{p}(s)\).
\end{lemma}

See the appendix (Section~\ref{sec:appendix-proof-details}) for details.

\begin{lemma}\label{lemma:k-order}
The \(K\)-order error \(\db{K}^{(0)}_0\) satisfies
\begin{equation}
  \left|\db{K - 1}^{(0)}_0 - \cdb{K - 1}^{(0)}_0\right| =
  \left|\db{K}^{(0)}_0\right| \leq \alpha(n) \mach^K
  \widetilde{p}(s).
\end{equation}
\end{lemma}

\begin{proof}
As in \eqref{matrix-de-casteljau}, we can express the compensated
de Casteljau algorithm as
\begin{equation}
\db{F}^{(k)} = U_{k + 1} \db{F}^{(k + 1)} + \ell_{F}^{(k)}
\Longrightarrow \db{F}^{(0)} = \sum_{k = 0}^{n - 1}
U_1 \cdots U_k \ell_F^{(k)} = \sum_{k = 0}^{n - 1}
\left[\sum_{j = 0}^k \ell_{F, j}^{(k)} B_{j, k}(s)\right].
\end{equation}
For the inexact equivalent of these things, first note that
\(\widehat{r} = (1 - s)(1 + \delta)\). Due to this,
we put the \(\widehat{r}\) term at the end of each update step to reduce
the amount of round-off:
\begin{align}
  \cdb{F}_j^{(k)} &=
  \widehat{\ell}_{F, j}^{(k)} \oplus
  \left(s \otimes \cdb{F}_{j + 1}^{(k + 1)}\right) \oplus
  \left(\widehat{r} \otimes \cdb{F}_j^{(k + 1)}\right) \\
&= (1 - s) \cdot \cdb{F}_j^{(k + 1)}(1 + \theta_3) +
  s \cdot \cdb{F}_{j + 1}^{(k + 1)}(1 + \theta_3) +
  \widehat{\ell}_{F, j}^{(k)} (1 + \theta_2) \\
\Longrightarrow \cdb{F}^{(k)} &=
  U_{k + 1} \cdb{F}^{(k + 1)}(1 + \theta_3) +
  \widehat{\ell}_{F}^{(k)} (1 + \theta_2) \\
\Longrightarrow \cdb{F}^{(0)} &=
  \sum_{k = 0}^{n - 1}
  U_1 \cdots U_k \widehat{\ell}_F^{(k)} (1 + \theta_{3k + 2})
  = \sum_{k = 0}^{n - 1}
  \left[\sum_{j = 0}^k \widehat{\ell}_{F, j}^{(k)} (1 + \theta_{3k + 2})
    B_{j, k}(s)\right].
\end{align}
Since
\begin{equation}
\db{F + 1}_0^{(0)} = \db{F}_0^{(0)} - \cdb{F}_0^{(0)} = \sum_{k = 0}^{n - 1}
\sum_{j = 0}^k \left(\ell_{F, j}^{(k)} -
\widehat{\ell}_{F, j}^{(k)} (1 + \theta_{3k + 2})\right) B_{j, k}(s)
\end{equation}
it's useful to put a bound on \(\ell_{F, j}^{(k)} -
\widehat{\ell}_{F, j}^{(k)} (1 + \theta_{3k + 2})\). Via
\begin{align}
\widehat{\ell}_{F, j}^{(k)} &= e_1 \oplus \cdots \oplus e_{5F - 2} \oplus
\left(\rho \otimes \cdb{F - 1}_j^{(k + 1)}\right) \\
&= e_1\left(1 + \theta_{5F - 2}\right) + \cdots +
e_{5F - 2}\left(1 + \theta_2\right) +
\rho \cdot \cdb{F - 1}_j^{(k + 1)} \left(1 + \theta_2\right)
\end{align}
we see that
\begin{equation}
\left|\ell_{F, j}^{(k)} -
\widehat{\ell}_{F, j}^{(k)} (1 + \theta_{3k + 2})\right| \leq
\gamma_{3k + 5F} \cdot \widetilde{\ell}_{F, j}^{(k)}
\Longrightarrow
\left|\db{F + 1}_0^{(0)}\right| \leq \sum_{k = 0}^{n - 1}
\gamma_{3k + 5F} \sum_{j = 0}^k \widetilde{\ell}_{F, j}^{(k)} B_{j, k}(s).
\end{equation}
Since Lemma~\ref{lemma:L-and-D-bounds} (in particular \eqref{L-sum-bound})
tells us that \(\sum_{j = 0}^k \widetilde{\ell}_{F, j}^{(k)} B_{j, k}(s)
= L_{F, k} = \mathcal{O}\left(\mach^F\right)\) ...
\end{proof}

\begin{theorem}
  If no underflow occurs, \(n \geq 2\) and \(s \in \left[0, 1\right]\)
  \begin{equation}
    \frac{\left|p(s) - \mathtt{CompDeCasteljau}(p, s, K)\right|}{
      \left|p(s)\right|} \leq \left[\mach + \mathcal{O}\left(\mach^2\right)
      \right] +
      \left[\alpha(n) \mach^K + \mathcal{O}\left(
      \mach^{K + 1}\right)\right] \cond{p, s}.
  \end{equation}
\end{theorem}
